ASHA 2023 Listening Effort Data Analysis

### 2023-10-29 ASHA 2023 Analysis/data_analysis.Rmd
updated pupil analysis code

@@

@@

# Model 1

@@

## Model 2

@@

```
## Model 3

@@

```
## Model 4

@@

>>>>

rm(als, control)

@@

The adjusted ICC for the unconditional growth model is .109. The unadjusted ICC for the conditional growth model is .107.

@@

In the next model, we are adding a fixed effect for speaker to test whether on average, pupil dilation is greater for the ALS speaker compared to the control speaker. A likelihood ratio test was also conducted to test whether model 1 provides a better fit to the data than the unconditional growth model.

@@

As indicated in the model output, when listening to the ALS speaker, on average, listeners had greater pupil dilation compared to when they were listening to the control speaker. The likelihood ratio test also indicates model 1 is a better fit to the data compared to the unconditional model (chi-sq(1) = 353.67, p <0.001).

In model two, the interaction between poly1 and speaker was added to the model as a fixed effect to test whether linear growth was greater for the ALS speaker compared to the control speaker. A likelihood ratio test was also conducted to test if model 2 provides a better fit to the data compared to model 1.

Note: The initial run of this model did not converge, so an optimizer control was added to the model. This optimizer will remain in each subsequent model.

@@

Based on the model output, the interaction between poly1 and speaker is significant, indicating that, across listeners, linear growth in pupil dilation while listening to the ALS speaker was significantly greater than when listeners were listening to the control speaker. The likelihood ratio test also indicated that model 2 provided a better fit to the data than model 1 (chi-sq(1) = 649.22, p <0.001).

In model three, the interaction between the quadratic growth term and speaker was added to the model. This tests whether the quadratic shape of pupil dilation is significantly different between the two speakers. A likelihood ratio test was also conducted to test whether model 3 provides a better fit to the data compared to model 2.

@@

```
Based on the model output, the interaction between the quadratic growth term and speaker improved the model fit (ch-sq(1) = 147.72, p<0.001). This indicates, on average, the shape of the listeners' pupil dilation is different while listening to the ALS speaker vs the control speaker.

In model for, we added the interaction between the cubic growth term and speaker. As with the other models, a likelihood ratio test was conducted to test whether model 4 provides a better fit to the data compared to model 3.

```{r, message = F, warning = F}

@@

```
Based on the model output, model 4 provides a better fit to the data compared to model 3 (chi-sq(1) = 9.25, P = 0.002). What is interesting is that the only fixed effects that are significant in this model are the main effect of the cubic growth term and the interaction between the cubic growth term and speaker, indicating the cubic pupil dilation growth is significantly different between these two speakers.

```{r, include = F, message = F, warning = F}

# Saving model 4 output

sjPlot::tab_model(m1, m2, m3, m4, file = "Final Model Outputs/pupil_dilation_model.html")

rm(m0, m1, m2, m3, m4, pupil_plot, pupil_roi, data, control_bobyqa, my_pal)
